---
title: "fin_paper"
output: github_document
---

```{r}
library(foreign)
library(tidyverse)
library(lubridate)


source("load_carls.R")
source("load_precinct_data.R")
tot_data = left_join(model_data,pres_summary,by=c("ELECTION_YEAR"="Year","Assembly"="ElectionType","DistrictID"="DISTRICT_ID"))
```

The format of the tidied dataset we want to model is this:

Every row is a (Year, Assembly, District) triple. This is a nice unit because it is the unit we want to run models on. However, none of our data is in this form. All of our data is in a (Year, Canidate) format. So in each dataset, before joining, we need to summarise the sort of covariates we want from that bit of canidate level data.

One of the major necessary transformations is that 


```{r load_economic_data}
#Import Nevada Data of Annual Median Household Income (Aggregated) 1984-2016
folder = "state-legislative-data/Economic Data/"
filename = "Nevada Annual Real Median Househod Income Data 1984-2017 BEA code MEHOINUSNVA672N.csv"
path = paste(folder, filename,sep="")
AggregateMedianIncomeData_Nevada = read_csv(path) %>%
  rename(MED_HOUSE_INCOME = MEHOINUSNVA672N) %>%
  mutate(YEAR = as.integer(format(DATE,"%Y")))

#Import Nevada Data on % Change of Annual Median Household Income (Marginal) 1984-2016
MedianIncomeData_Nevada = read_csv("state-legislative-data/Economic Data/Nevada Percent Change in Annual Real Median Househod Income Data 1984-2016.csv") %>%
  rename(MHI_Change_LOCAL = MEHOINUSNVA672N_PCH) %>%
  mutate(YEAR = as.integer(format(DATE,"%Y")))

#Import Nevada's Annual Unemployment Rate Data 1976-2016 
Unemployment_Nevada <- read_csv("state-legislative-data/Economic data/Nevada Annual Unemployment Rate 1976-2016.csv") %>% 
  rename(Unemployment_LOCAL = NVUR) %>%
  mutate(YEAR = as.integer(format(DATE,"%Y")))

#Import National Data on % Change of Annual Median houseHold Income 1984-2016
MedianIncomeData_US <- read_csv("state-legislative-data/Economic data/US National-Level Percent Change in Annual Real Median Household Income 1984-2016.csv") %>% 
  rename(MHI_Change_US = MEHOINUSA672N_PCH) %>%
  mutate(YEAR = as.integer(format(DATE,"%Y")))

#Import National Data on Percent Change of Annual Real GDP/Capita 1948-2016
GDPperCapita_US <- read_csv("state-legislative-data/Economic data/US National-Level Percent Change in Annual Real GDP per Capita 1948-2016.csv") %>% 
  rename(GDPperCapita_CHANGE_US = A939RX0Q048SBEA_PCH) %>%
  mutate(YEAR = as.integer(format(DATE,"%Y")))

#Join the Previous 4 Data Fragments Together into one Frame  (labeled Economic Data)
economic_data <- GDPperCapita_US %>%
  left_join(MedianIncomeData_Nevada, by=c("YEAR" = "YEAR")) %>%
  left_join(MedianIncomeData_US, by=c("YEAR" = "YEAR")) %>%
  left_join(Unemployment_Nevada, by=c("YEAR" = "YEAR")) %>%
  select(3, 2, 5, 7, 9)
```

```{r join econ with voting data}

with_economic = left_join(tot_data, economic_data ,by=c("ELECTION_YEAR"="YEAR"))
#economic_data %>% left_join(tot_data, by=c("YEAR"="ELECTION_YEAR"))
```

```{r graph}
#Calculate percentage of house captured by Democrats during a Given Year's General Election
percent_seat_data <- tot_data %>%
  group_by(ELECTION_YEAR, Assembly) %>%
  summarize(perc_seats_dem = sum(wining_party=="DEM")/n()) %>%
  left_join(economic_data,by=c("ELECTION_YEAR"="YEAR"))


#Tidy df so that we can graph data on a 2 dimensional plane 
#Exclude unemployment rate (it stretches te y-axis scale too far to capture insights on
                          # the correlating patterns between voting and economic data)
with_economic_Tidied1 <- percent_seat_data %>% gather(Type,Val,-ELECTION_YEAR, -Assembly, 
                                                  -Unemployment_LOCAL)
#Visualize 
graph1 <- ggplot(with_economic_Tidied1,aes(x=ELECTION_YEAR, y=Val, col=Type)) + 
  geom_line() + 
  facet_wrap(~Assembly)

#Tidy a separate df by gathering just Unemployment_Local  and perc_seats_dem under Type
with_economic_Tidied2 <- percent_seat_data %>% gather(key = Type, value = Percent, 
                                                  Unemployment_LOCAL, perc_seats_dem)
```

```{r}
beforeyear = 2014

train_data = use_data %>%
  filter(ELECTION_YEAR < beforeyear) %>%
  filter(ELECTION_YEAR > 1970) %>% # we don't have district numbers for 1968 and 1970! (this should be fixable)
  group_by(DISTRICT_NUM,ELECTION_YEAR) %>%
  mutate(percent_vote = CANIDATE_VOTE_TOTAL/sum(CANIDATE_VOTE_TOTAL)) %>%
  ungroup()

pred_perc = lm(percent_vote ~ INCUMBENCY_DUMMY,data=train_data)
```


```{r}
senate_data %>%
  filter(ELECTION_YEAR >= 2012) %>%
  select(ALT_DISTRICT_NUM,DISTRICT_NAME) %>%
  unique()
```

Road citation:

Gary King; Bradley Palmquist; Greg Adams; Micah Altman;
Kenneth Benoit; Claudine Gay; Jeffrey B. Lewis; Russ Mayer;
and Eric Reinhardt. 1997. "The Record of American Democracy, 1984-1990," Harvard University, Cambridge, MA [producer],
Ann Arbor, MI: ICPSR [distributor].