---
title: "fin_paper"
output: github_document
---

```{r}
library(foreign)
library(tidyverse)
library(lubridate)


source("load_carls.R")
source("load_precinct_data.R")
source("load_econ_data.R")

```

```{r district_level_data}
pres_summary = prec_data_2004_2014 %>% 
  group_by(officename,Year,SENATE_OR_HOUSE,DIST_NAME,DIST_NUM,Selection) %>%
  summarise(vote_count = sum(Votes),
            Party = get_pres_gov_party(Selection[1])) %>%
  summarise(vote_dem = sum(ifelse(Party=="DEM",vote_count,0)),
            total_vote = sum(vote_count)) %>%
  ungroup() %>%
  gather(key=value_type,value=vote_value,vote_dem:total_vote)%>%
  unite(RaceValue,officename,value_type) %>%
  spread(RaceValue,vote_value)

named_data = use_data %>%
  left_join(filter(pres_summary,!is.na(DIST_NAME)),
            by=c("ELECTION_YEAR"="Year","Assembly"="SENATE_OR_HOUSE","DISTRICT_NUM"="DIST_NUM","DISTRICT_NAME"="DIST_NAME"))

unnamed_data = use_data %>%
  left_join(filter(pres_summary,is.na(DIST_NAME)),
            by=c("ELECTION_YEAR"="Year","Assembly"="SENATE_OR_HOUSE","DISTRICT_NUM"="DIST_NUM")) %>%
  select(-DIST_NAME)

district_level_data = rbind(unnamed_data,named_data) %>%
  mutate(party_val = ifelse(wining_party == "DEM",1,
                            ifelse(wining_party == "REP",-1,0)))
```

The format of the tidied dataset we want to model is this:

Every row is a (Year, Assembly, District) triple. This is a nice unit because it is the unit we want to run models on. However, none of our data is in this form. All of our data is in a (Year, Canidate) format. So in each dataset, before joining, we need to summarise the sort of covariates we want from that bit of canidate level data.

```{r state_level_data}
pres_party_in_power_data = read_csv("state-legislative-data/pres_in_power.csv")

state_level_data = economic_data %>%
  left_join(pres_party_in_power_data,by=c("YEAR"="year")) %>%
  
  # spread party_in_power variable over all the years they are in power, not just every 4 years
  group_by((YEAR-1) %/% 4) %>%
  mutate(pres_party_in_power = paste(ifelse(is.na(pres_party_in_power),"",pres_party_in_power),collapse="")) %>%
  ungroup() %>%
  
  #calculate midterm and presidentail election penalties
  mutate(is_midterm = (YEAR %% 4) == 2,
         is_pres_election = (YEAR %% 4) == 0,
         pres_party_num = ifelse(pres_party_in_power=="DEM",1,
                                 ifelse(pres_party_in_power=="REP",-1,0)),
         midterm_penalty = pres_party_num * is_midterm,
         pres_elect_penalty = pres_party_num * is_pres_election) %>%
  
  select(YEAR:Unemployment_LOCAL,midterm_penalty,pres_elect_penalty)
```

```{r assembly_level_summary}
year_summary = precinct_joined %>%
  group_by(Assembly,ELECTION_YEAR) %>%
  summarise(senate_composition = sum(party_val))
```

```{r}
beforeyear = 2014
  
single_level_model_data = precinct_joined %>%
  left_join(year_summary ,by=c("ELECTION_YEAR","Assembly")) %>%
  left_join(state_level_data ,by=c("ELECTION_YEAR"="YEAR"))

train_data = single_level_model_data %>%
  filter(ELECTION_YEAR < beforeyear) %>%
  filter(ELECTION_YEAR > 1970) %>% # we don't have district numbers for 1968 and 1970! (this should be fixable)
  group_by(DISTRICT_NUM,DISTRICT_NAME,ELECTION_YEAR) %>%
  ungroup()

pred_perc = lm(senate_composition ~ 
                 INCUMBENCY_DUMMY + 
                 
                 ,data=train_data)
```

```{r}
senate_data %>%
  filter(ELECTION_YEAR >= 2012) %>%
  select(ALT_DISTRICT_NUM,DISTRICT_NAME) %>%
  unique()
```

```{r}
#Calculate percentage of house captured by Democrats during a Given Year's General Election
percent_seat_data <- tot_data %>%
  group_by(ELECTION_YEAR, Assembly) %>%
  summarize(perc_seats_dem = sum(wining_party=="DEM")/n()) %>%
  left_join(economic_data,by=c("ELECTION_YEAR"="YEAR"))


#Tidy df so that we can graph data on a 2 dimensional plane 
#Exclude unemployment rate (it stretches te y-axis scale too far to capture insights on
# the correlating patterns between voting and economic data)
with_economic_Tidied1 <- percent_seat_data %>% gather(Type,Val,-ELECTION_YEAR, -Assembly, 
                                                      -Unemployment_LOCAL)
#Visualize 
graph1 <- ggplot(with_economic_Tidied1,aes(x=ELECTION_YEAR, y=Val, col=Type)) + 
  geom_line() + 
  facet_wrap(~Assembly)

#Tidy a separate df by gathering just Unemployment_Local  and perc_seats_dem under Type
with_economic_Tidied2 <- percent_seat_data %>% gather(key = Type, value = Percent, 
                                                      Unemployment_LOCAL, perc_seats_dem)
```

Road citation:

Gary King; Bradley Palmquist; Greg Adams; Micah Altman;
Kenneth Benoit; Claudine Gay; Jeffrey B. Lewis; Russ Mayer;
and Eric Reinhardt. 1997. "The Record of American Democracy, 1984-1990," Harvard University, Cambridge, MA [producer],
Ann Arbor, MI: ICPSR [distributor].